Local Spectral Clustering of Density Upper Level Sets

### 2019-05-14 arxiv/arxiv_local_density_clustering_proofs.tex
updated proofs, detailed edits of arxiv version. Copied/pasted neurips version

\usepackage{microtype}
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
% \usepackage{hyperref}
\usepackage{bm}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{xr-hyper}
\usepackage{natbib}
\usepackage{fullpage}

\let\pprspace\relax
\let\ppr\relax
\externaldocument{arxiv_local_density_clustering}

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\left\{#1\right\}_{n \in \mathbb{N}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\Log}{\mathrm{Log}}
\newcommand{\TV}{\mathrm{TV}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Id}{\mathrm{Id}}

@@

%%% Vectors
\newcommand{\pbf}{\mathbf{p}}
\newcommand{\qbf}{\mathbf{q}}
\newcommand{\ebf}[1]{\mathbf{e}_{#1}}
\newcommand{\pibf}{\bm{\pi}}
\newcommand{\rhobf}{\bm{\rho}}
\newcommand{\Deltabf}{\bm{\Delta}}
\newcommand{\deltabf}{\bm{\delta}}
\newcommand{\zbf}{\mathbf{z}}

@@

%%% Matrices
\newcommand{\Abf}{\mathbf{A}}
\newcommand{\Xbf}{\mathbf{X}}
\newcommand{\Wbf}{\mathbf{W}}
\newcommand{\Lbf}{\mathbf{L}}
\newcommand{\Dbf}{\mathbf{D}}
\newcommand{\Ibf}[1]{\mathbf{I}_{#1}}

@@

%\newcommand{\theHalgorithm}{\arabic{algorithm}}

%\icmltitlerunning{Local clustering of density upper level sets}

@@

%\twocolumn[
%\icmltitle{Supplement to ``Local clustering of density upper level sets''}

%\icmlsetsymbol{equal}{*}

%\begin{icmlauthorlist}
%\icmlauthor{Alden Green}{cmu}
%\icmlauthor{Sivaraman Balakrishnan}{cmu}
%\icmlauthor{Ryan Tibshirani}{cmu}
%\end{icmlauthorlist}

%\icmlaffiliation{cmu}{Department of Statistics and Data Science, Carnegie Mellon University, Pittsburgh PA, USA}

%\icmlcorrespondingauthor{Alden Green}{ajgreen@andrew.cmu.edu}

%\icmlkeywords{local clustering}

%\vskip 0.3in
%]

%\printAffiliationsAndNotice{} % otherwise use the standard text.

@@

In this supplement, we present proofs for ``Local Clustering of Density Upper Level Sets''. Sections \ref{sec: volume_estimates} - \ref{sec: proof_of_theorem_1} detail the proof for Theorem \ref{thm: conductance_upper_bound}. \ref{sec: mixing_time_on_graphs} develops a bound of the form of \eqref{eqn: average_conductance}, which we recall links the conductance function to mixing time; this will be necessary for both Theorems \ref{thm: inverse_mixing_time_lower_bound_nonconvex} and \ref{thm: inverse_mixing_time_lower_bound}. \ref{sec: non_convex_conductance_function_and_local_spread} and \ref{sec: inverse_mixing_time_lower_bound_nonconvex} give the proof of Theorem \ref{thm: inverse_mixing_time_lower_bound_nonconvex}, while \ref{sec: convex_population_conductance_function}- \ref{sec: inverse_mixing_time_lower_bound} give the proof of Theorem \ref{thm: inverse_mixing_time_lower_bound}. \ref{sec: concentration} gives some general concentration results used throughout, before we finish with the proof of Theorem \ref{thm: consistent_recovery_of_density_clusters} in \ref{sec: proof_of_consistent_cluster_recovery}.

@@

	We first consider the term $\Sigma_k$. $\Cset$ has finite diameter by \ref{asmp: cluster_diameter}. Letting $\overline{\Cset}$ be the closure of $\Cset$, we observe that $\overline{\Csig} = \overline{\Cset} + \sigma B$, and moreover for any $\delta > 0$, $\nu(\overline{\Csig} + \delta B) = \nu(\Csig + \delta B)$ (as $\partial(\Csig + \delta B)$ is Lipschitz and therefore has measure zero). As a result, for each $t_{i,k}, i = 1, \ldots,k$ we may apply Lemma \ref{lem: expansion_volume} to $\overline{\Cset}$ and obtain
	\begin{equation}
	\label{eqn: slice_volume_bound}
	\nu_{i,k} = \nu(\Csig + rt_{i,k}B) \leq \nu(\Csig)\left(1 + \frac{rt_{i,k}}{\sigma}\right)^d
	\end{equation}
	which in turn gives
	\begin{align}
	\Sigma_k & \leq c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ rt_{i,k}}{\sigma}\right)^d \biggl( (t_{i,k})^{\gamma} - (t_{i-1,k})^{\gamma}\biggr) \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}).~~~~~~~~~~~~~~ (u_{i,k} = t_{i,k}^{\gamma}) \label{eqn: Sigmak_riemann_sum}
	\end{align}
	\eqref{eqn: Sigmak_riemann_sum} is a Riemann sum, and taking the limit as $k \to \infty$ we obtain
	\begin{align}
	\lim_{k \to \infty} c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}) & = c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{r u^{1/\gamma}}{\sigma}\right)^{d} du \nonumber \\
	& \overset{(i)}{\leq} c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{2 d r u^{1/\gamma}}{\sigma}\right) du \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \left(1 + \gamma \frac{2 d r}{(\gamma + 1)\sigma}\right). \label{eqn: Sigmak_integral}
	\end{align}
	where $(i)$ follows from Lemma \ref{lem: Taylor_series} in light of the fact $r \leq \sigma / 2d$.

@@

For $N \in \mathbb{N}$ and a collection $V = \set{v_1,\ldots,v_N}$ of $N$ vertices, take $G = (V,E)$ to be an undirected and unweighted graph, with associated adjacency matrix $\Abf$, degree matrix $\Dbf$, and random walk matrix $\Wbf$. We consider the stationary distribution $\pi$ to be a mapping from $V$ to $[0,1]$, with $\pi(u) := \frac{\Dbf_{uu}}{\vol(V; G)}$ for $u \in V$. For $v \in V$, let $q: V \times V \to [0,1]$ represent the $m$-step transition probability of the random walk over $G$ originating from $v$, $q^{(m)}(v,u) = e_v\Wbf^m e_u$. The relative pointwise mixing time over $G$ is then
\begin{equation*}
\tau_{\infty}(G) = \min\set{m: \forall u,v \in V, \frac{\abs{q^{(m)}(v,u) - \pi(u)}}{\pi(u)} \leq 1/4}. 
\end{equation*}

Two key quantities relate the mixing time to the expansion of subsets $S$ of $V$. The \emph{local spread} is defined as
\begin{equation*}
s(G) := \frac{9}{10} \cdot \min_{v \in V} \set{\deg(v; G)} \cdot \min_{v \in V} \set{\pi(v)}
\end{equation*}
For $f: V \to \Reals$, and $S \subseteq V$, write $f(S) := \sum_{u \in S} f(u)$. We collect some necessary facts about the local spread in Lemma \ref{lem: local_spread_G}.
\begin{lemma}
	~

	\label{lem: local_spread_G}
	\begin{itemize}
		\item If $\pi(S) \leq s(G)$, then for every $u \in S$, $q^{(1)}(u,S^c) \geq \frac{1}{10}$.
		\item For any $v, u \in V$, and $m \in \mathbb{N}, m \geq 0$,
		\begin{equation*}
		q^{(m)}(v,u) \leq \frac{\min_{v \in V} \set{\pi(v)}}{s(G)}.
		\end{equation*}
	\end{itemize}

\end{lemma}

\begin{proof}
	If $\pi(S) \leq s(G)$, divide both sides by $\min_{v \in V} \set{\pi(v)}$ to obtain $\abs{S} \leq 9\min_{v \in V} \set{\deg(v; G)}/10$,
	which implies $q^{(1)}(v,S^c) \geq 1/10$ for all $v \in S$. This implies the first statement.

	The second statement follows from the fact $q^{(m)}(v,u) \leq 1/\min_{v \in V} \set{\deg(v; G)}$ for any $m$.
\end{proof}

The local spread facilitates conversion between $\tau_{\infty}(G)$ and the more easily manageable \emph{total variation} mixing time, given by
\begin{equation*}
\tau_1(G) = \min\biggl\{m: \forall v \in V, \norm{\rho^{(m)}(v,\cdot) - \pi}_{TV} \leq 1/4 \biggr\}
\end{equation*}
where $\rho^{(m)}(u,v) = \sum_{l = 1}^{m} p^{(l)}(u,v)$ 
and $\norm{\rho(v,\cdot) - \pi}_{TV} = \sum_{u \in V}\abs{\rho(u,v) - \pi(v)}$ is the total variation norm between distributions $\rho(v,\cdot)$ and $\pi$. 
\begin{lemma}
	\label{lem: tv_mixing_to_pointwise_mixing1}
	\begin{equation*}
	\tau_{\infty}(G) \leq 2752 \cdot \tau_1(G) \log \left(4 \max\left\{1, \frac{1}{s(G)}\right\}\right)
	\end{equation*}
\end{lemma}
\begin{proof}
	Masking dependence on the starting vertex $v$ for the moment, let
	\begin{equation*}
	\Delta^{(m)}(u) = q^{(m)}(v,u) - \pi(u), ~~ \delta^{(m)}(u) = \frac{\Delta^{(m)}(u)}{\pi(u)}
	\end{equation*}
	and denote $\Deltabf^{(m)} = \bigl(\Delta^{(m)}(u)\bigr)_{u \in V}$, $\deltabf^{(m)} = \bigl(\delta^{(m)}(u)\bigr)_{u \in V}$. For a vector $\Deltabf = (\Delta_u)_{u \in V}$, the $L^{p}(\pi)$ norm is given by
	\begin{equation*}
	\norm{\Deltabf}_{L^p(\pi)} = \left(\sum_{u \in V} \left(\Delta_u\right)^{p} \pi(u) \right)^{1/p}
	\end{equation*}
	To go between the $L^{\infty}(\pi)$ and $L^{1}(\pi)$ norms, we capitalize on a clever argument developed in \cite{morris2005}. For any $v,u \in V$, by the time-reversibility of the graph random walk,
	\begin{align}
	\abs{\frac{q^{(2m)}(v,u) - \pi(u)}{\pi(u)}} & = \abs{\sum_{y \in V} \pi(y) \left( \frac{q^{(m)}(v,y) - \pi(y)}{\pi(y)}\right)\left(\frac{q^{(m)}(u,y) - \pi(y)}{\pi(y)} \right)} \nonumber \\
	& \leq \norm{(\deltabf^{(m)})}_{L^{1}(\pi)} \norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)}, \label{eqn: infty_norm_to_one_norm}
	\end{align}
	with the second line following by Holder's inequality.

	Now, we upper bound the second factor on the right hand side by observing
	\begin{align}
	\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} & \leq \max\left\{1, \max_{u \in V} \frac{q^{(m)}(v,u)}{\pi(u)} \right\} \nonumber \\
	& \overset{(i)}{\leq} \max\left\{1, \frac{1}{s(G)}\right\} \label{eqn: ub_infinity_norm}
	\end{align}
	where $(i)$ follows from Lemma \ref{lem: local_spread_G}.

	Set $m_0 = 1376 \tau_1(G) $, which we will show implies $\norm{q^{(m)}(v,\cdot) - \pi}_{TV} \leq \frac{1}{4}$. Now, we leverage the following well-known fact (see pg. 1 of \cite{montenegro2002}): for any $\epsilon > 0$, if $m \geq m_0 \cdot \log(1/\epsilon)$ then for any $v \in V$,
	\begin{equation}
	\label{eqn: rapid_mixing}
	\norm{q^{(m)}(v,\cdot) - \pi}_{TV} \leq \epsilon.
	\end{equation}
	But $\norm{q^{(m)}(v,\cdot) - \pi}_{TV}$ is exactly $\norm{\deltabf^{(m)}}_{L^{1}(\pi)}$. Letting
	\begin{equation*}
	m_1 = m_0 \cdot \log \left(4 \max\left\{1, \frac{1}{s(G)}\right\} \right)
	\end{equation*} 
	by \eqref{eqn: infty_norm_to_one_norm}, \eqref{eqn: ub_infinity_norm}, and \eqref{eqn: rapid_mixing}, $\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} \leq 1/4$ for all $m \geq 2 m_1$.  Recall that
	\begin{equation*}
	\norm{(\deltabf^{(m)})}_{L^{\infty}(\pi)} = \sup_{u \in V}\left\{ \frac{\abs{q^{(m)}(v,u) - \pi(u)}}{\pi(u)} \right\}.
	\end{equation*}
	and since none of the above depended on a specific choice of $v$, the supremum can be taken over all starting vertices $v$ as well. Thus for all $m \geq 2m_1$, for any $v,u \in V$, $\abs{q^{m}(v,u) - \pi(u)} \leq \frac{\pi(u)}{4}$.

	Finally, it is known (pg.15 of \citep{montenegro2002}) that
	\begin{equation*}
	\norm{\rho^{(m)}(v,\cdot) - \pi}_{TV} \leq \frac{1}{4} \Longrightarrow  \norm{q^{(1376m)}(v,\cdot) - \pi}_{TV} \leq \frac{1}{4} 
	\end{equation*}
	and so $m_0 = 1376 \tau_1(G)$ implies $\norm{q^{(m)}(v,\cdot) - \pi}_{TV} \leq \frac{1}{4}$, completing the proof.
\end{proof}

The second key quantity is the \emph{conductance function}. For $\min_{u \in V}\set{\pi(u)} \leq t \leq 1/2$, the conductance function is defined as
\begin{equation}
\label{eqn: conductance_function}
\Phi(t; G) := \min_{\substack{S \subseteq V, \\ 0 < \pibf(S) \leq t} } \Phi(S; G)
\end{equation}
where $\Phi(S; G)$ is the normalized cut of $S$ in $G$ given by \eqref{eqn: norm_cut}, and in the minimization $S$ may not be the empty set.

Lemma \ref{lem: expansion_to_tv_distance} leverages the conductance function and local spread to produce an upper bound on the total variation distance between $\rhobf_{v}^{(m)}$ and $\pibf.$
\begin{lemma}
	\label{lem: expansion_to_tv_distance}
	If $D_{\min} > 10$, for any $v \in V$:
	\begin{equation*}
	\norm{\rhobf_v^{(m)} - \pibf}_{TV} \leq \max\left\{ \frac{1}{4}, \frac{1}{10} +  \frac{70}{m}\left(\frac{20}{9} + \int_{t = s'(G)}^{1/2} \frac{4}{t \Phi^2(t; G)} dt\right) \right\}
	\end{equation*}
	where $s'(G) = s(G) / 9$.
\end{lemma}

To prove Lemma \ref{lem: expansion_to_tv_distance} we first introduce a generalization of $\Phi(t;G) \cdot \Phi(t;G)$ known as a blocking conductance function.
\footnote{For more details, see \cite{montenegro2002}}

\begin{definition}[Blocking Conductance Function of \cite{montenegro2002}]
	\label{def: blocking_conductance_function}
	For $t_0 \geq \pi_{\min}$, a function $\phi(t; G): [t_0, 1/2] \to [0,1]$ is a \emph{blocking conductance function} if for all $S \subset V$ with $\pibf(S) = t \in [t_0, 1/2]$, either of the following hold: 
	\begin{enumerate}
		\item \emph{Exterior inequality.} For all $y \in \left[\frac{1}{2}t, t\right]: \phi_{int}(S) \geq \phi(\max\{t_0,y\})$
		\item \emph{Interior inequality.} For all $y  \in \left[t, \frac{3}{2}t\right]: \phi_{ext}(S) \geq \phi(\max\{y,1 - y\})$.
	\end{enumerate}
	where $\phi_{int}$ and $\phi_{ext}$ are defined respectively as
	\begin{align*}
	\phi_{int}(S) & = \sup_{\lambda \leq \pibf(S)} \min_{\substack{B \subset S \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\pibf(S) \pibf(S^c)\right]^2} \\
	\phi_{ext}(S) & = \sup_{\lambda \leq \pibf(S)} \min_{\substack{B \subset S^c \\ \pibf(B) \leq \lambda} } \frac{\lambda \cut(S \setminus B, S^c; G)}{\vol(V; G) \left[\pibf(S) \pibf(S^c)\right]^2}
	\end{align*}
\end{definition}

>>>>

\usepackage{microtype}
\usepackage{graphicx}
\usepackage[export]{adjustbox}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts, amsthm, amssymb}
\usepackage[parfill]{parskip}
\usepackage{enumerate}
\usepackage[shortlabels]{enumitem}
% \usepackage{hyperref}
\usepackage{bm}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue,linkcolor=blue]{hyperref}
\usepackage{xr-hyper}
\usepackage{natbib}
\usepackage{fullpage}
\usepackage{mathtools}

\let\pprspace\relax
\let\ppr\relax
\let\Cset\relax
\let\Csig\relax

\externaldocument{arxiv_local_density_clustering}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}

\newcommand{\diam}{\mathrm{diam}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\seq}[1]{\left\{#1\right\}_{n \in \mathbb{N}}}
\newcommand{\defeq}{\overset{\mathrm{def}}{=}}
% \newcommand{\abs}[1]{\left \lvert #1 \right \rvert}
\newcommand{\vol}{\mathrm{vol}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Reals}{\mathbb{R}}
\newcommand{\Rd}{\Reals^d}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\1}{\mathbf{1}}
\newcommand{\var}{\mathrm{Var}}
\newcommand{\Err}{\mathrm{Err}}
\newcommand{\Log}{\mathrm{Log}}
\newcommand{\TV}{\mathrm{TV}}
\newcommand{\dist}{\mathrm{dist}}
\newcommand{\Id}{\mathrm{Id}}

@@

%%% Vectors
\newcommand{\pbf}{p} % no bold font
\newcommand{\qbf}{\mathbf{q}}
\newcommand{\ebf}[1]{\mathbf{e}_{#1}}
\newcommand{\pibf}{\pi} % no bold font
\newcommand{\rhobf}{\bm{\rho}}
\newcommand{\Deltabf}{\bm{\Delta}}
\newcommand{\deltabf}{\bm{\delta}}
\newcommand{\zbf}{\mathbf{z}}

@@

%%% Matrices
\newcommand{\Abf}{\mathbf{A}}
\newcommand{\Xbf}{X} % no bold font
\newcommand{\Wbf}{\mathbf{W}}
\newcommand{\Lbf}{\mathbf{L}}
\newcommand{\Dbf}{\mathbf{D}}
\newcommand{\Ibf}[1]{\mathbf{I}_{#1}}

@@

\theoremstyle{remark}
\newtheorem{remark}{Remark}

@@

@@

In this supplement, we present proofs for ``Local Clustering of Density Upper Level Sets''. Sections \ref{sec: volume_estimates} - \ref{sec: proof_of_theorem_1} detail the proof for Theorem \ref{thm: conductance_upper_bound}. In Sections \ref{sec: mixing_time_on_graphs} and \ref{sec: proof_of_proposition_tv_mixing_time}, we establish a link between the conductance and local spread, and mixing time; this will be necessary for Theorem \ref{thm: mixing_time_upper_bound}. The proof of Theorem \ref{thm: mixing_time_upper_bound} is then carried out in Sections \ref{sec: population_conductance_function} - \ref{sec: proof_of_theorem_mixing_time}. \ref{sec: concentration} gives some general concentration results we use throughout, before we finish with the proofs of Theorems \ref{thm: misclassification_rate} and \ref{thm: consistent_recovery_of_density_clusters}, and Corollary \ref{cor: appr}, in Sections \ref{sec: proof_of_misclassification_rate} - \ref{sec: linalg}.

@@

	% AJG 5/14/19: This argument makes sense, correct?
	We first consider the term $\Sigma_k$. $\Cset$ has finite diameter by \ref{asmp: bounded_density}, as otherwise $\int_{\Csig} f(x) dx = \infty$. Letting $\overline{\Cset}$ be the closure of $\Cset$, we observe that $\overline{\Csig} = \overline{\Cset} + \sigma B$, and moreover for any $\delta > 0$, $\nu(\overline{\Csig} + \delta B) = \nu(\Csig + \delta B)$ (as $\partial(\Csig + \delta B)$ is Lipschitz and therefore has measure zero). As a result, for each $t_{i,k}, i = 1, \ldots,k$ we may apply Lemma \ref{lem: expansion_volume} to $\overline{\Cset}$ and obtain
	\begin{equation}
	\label{eqn: slice_volume_bound}
	\nu_{i,k} = \nu(\Csig + rt_{i,k}B) \leq \nu(\Csig)\left(1 + \frac{rt_{i,k}}{\sigma}\right)^d
	\end{equation}
	which in turn gives
	\begin{align}
	\Sigma_k & \leq c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ rt_{i,k}}{\sigma}\right)^d \biggl( (t_{i,k})^{\gamma} - (t_{i-1,k})^{\gamma}\biggr) \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}).~~~~~~~~~~~~~~ (u_{i,k} = t_{i,k}^{\gamma}) \label{eqn: Sigmak_riemann_sum}
	\end{align}
	\eqref{eqn: Sigmak_riemann_sum} is a Riemann sum, and taking the limit as $k \to \infty$ we obtain
	\begin{align}
	\lim_{k \to \infty} c_0\nu(\Csig) r^\gamma \sum_{i = 1}^{k} \left(1 + \frac{ru_{i,k}^{1/\gamma}}{\sigma}\right)^d ( u_{i,k} - u_{i,k-1}) & = c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{r u^{1/\gamma}}{\sigma}\right)^{d} du \nonumber \\
	& \overset{(i)}{\leq} c_0\nu(\Csig) r^\gamma \int_{0}^{1} \left(1 + \frac{2 d r u^{1/\gamma}}{\sigma}\right) du \nonumber \\
	& = c_0\nu(\Csig) r^\gamma \left(1 + \gamma \frac{2 d r}{(\gamma + 1)\sigma}\right). \label{eqn: Sigmak_integral}
	\end{align}
	where $(i)$ follows from Lemma \ref{lem: Taylor_series} in light of the fact $r \leq \sigma / 2d$.

@@

We recall some notation from the main text. For an undirected graph $G = (V,E)$, the lazy random walk over $G$ is the Markov chain with transition probabilities given by $\Wbf := \frac{\Ibf + \Dbf^{-1} \Abf}{2}$ and stationary distribution $\pi$.  Denote the $m$-step probability distribution of this random walk originating from a particular $v \in V$ as $q^{(m)}: V \times V \to [0,1]$, $q^{(m)}(v,u) = e_v \Wbf^m e_u$. An important intermediate quantity used to bound the relative pointwise mixing time $\tau_{\infty}(G)$ is the total variation distance between the distributions $q_v^{(m)} := q^{(m)}(v,\cdot)$ and $\pi$,
\begin{equation*}
\norm{q_v^{(t + 1)} - \pi}_{TV} = \sum_{u \in V} \abs{q_v^{(m)}(u) - \pi(u)}
\end{equation*}
We recall the \emph{degree, cut} and \emph{volume} functionals over a graph. For $u \in V$, $S \subseteq V$, 
\begin{equation*}
\cut(A;G) = \sum_{u \in A} \sum_{v \in A^c} \1((u,v) \in E), \quad \deg(u;G) = \sum_{v \in V} \1((u,v) \in E), \quad \vol(A;G) = \sum_{u \in A} \deg(u;G)
\end{equation*}

The \emph{local spread} is defined as
\begin{equation*}
s(G) := \frac{9}{10} \cdot \min_{u \in V} \set{\deg(u; G)} \cdot \min_{u \in V} \set{\pi(v)}
\end{equation*}

Letting the \emph{normalized cut} $\Phi(S;G)$ be defined as
\begin{equation*}
\Phi(S;G) = \frac{\cut(S, S^c; G)}{\min\set{\vol(S; G),\vol(S^c; G)}},
\end{equation*}
the \emph{conductance} is
\begin{equation*}
\Phi(G) = \min_{S \subseteq V} \Phi(S;G).
\end{equation*}

Proposition \ref{prop: tv_mixing_time} relates these geometric quantities to the TV distance between $q_v^{(t)}$ and stationary distribution $\pi$.
\begin{proposition}
	\label{prop: tv_mixing_time}
	For any $v \in V$, and any $0 < a \leq 1$,
	\begin{equation*}
	\norm{q_v^{(t + 3)} - \pi}_{TV} \leq \set{as(G), \frac{1}{8} + \frac{1}{20} + \frac{1}{2 \min_{u \in V} \deg(u;G)}} + \left(\frac{1}{1 - 2as(G)} + \frac{1}{2 as(G)}\right) \left(1 - \frac{\Phi^2(G)}{2}\right)^t
	\end{equation*}
\end{proposition}

As we will see, Proposition \ref{prop: tv_mixing_time} is an essential step to providing an upper bound on the uniform mixing time. We justify this statement next, before moving on to proving Proposition \ref{prop: tv_mixing_time} in the following subsection.

\paragraph{Uniform mixing time.}
Consider the uniform distance \footnote{Note $d_{\textrm{unif}}$ is not a formally a distance as it is not symmetric.} between $q_v^{(t)}$ and $\pi$, given by
\begin{equation*}
d_{\textrm{unif}}(q_v^{(t)},\pi) = \max_{u \in V} \set{\frac{\pi(u) - q_v^{(t)}(u)}{\pi(u)}}.
\end{equation*}

\begin{lemma}
	\label{lem: tv_to_uniform_distance}
	Let $\norm{q_v^{(t)} - \pi}_{TV} \leq \frac{1}{14} \max \set{1, \frac{1}{s(G)}}$. Then,
	\begin{equation*}
	d_{\mathrm{unif}}(q_v^{(t + 3)},\pi) \leq \frac{1}{4}
	\end{equation*}
\end{lemma}
\begin{proof}
	Fix $u \in V$ and let $m \geq t + 1$ be arbitrary. The stationarity of $\pi$ gives
	\begin{align}
	\frac{\pi(u) - q_v^{m}(u)}{\pi(u)} & = \sum_{y \in V} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \nonumber \\
	& \overset{(i)}{=} \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{\pi(u)} \left(\frac{1}{2} - \pi(u)\right) \nonumber \\
	& \leq \sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) + \frac{\pi(u) - q^{(m - 1)}(v,u)}{2 \pi(u)} \label{eqn: tv_to_uniform_distance_1}
	\end{align}
	where $(i)$ follows from $q^{(1)}(u,u) = \frac{1}{2}$.

	Then 
	\begin{align*}
	\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) & \leq \norm{q_v^{(m-1)} - \pi}_{TV} \max_{y \neq u} \abs{\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}} \\
	& \leq \norm{q_v^{(m-1)} - \pi}_{TV} \max \set{1, \max_{y \neq u}\set{\frac{q^{(1)}(y,u)}{\pi(u)}}} \\
	& \leq \norm{q_v^{(m-1)} - \pi}_{TV} \max \set{1, \frac{1}{s(G)}} \\
	\end{align*}
	since for $y \neq u$, $q^{(1)}(y,u) \leq 1/\left(2 \min_{u \in V} \deg(u; G)\right)$. As $m - 1 \geq t$, it is well known \cite{lovasz1990} that the laziness  of the random walk guarantees  $\norm{q_v^{(m - 1)} - \pi}_{TV} \leq \norm{q_v^{(t)} - \pi}_{TV}$, and therefore
	\begin{equation*}
	\sum_{y \neq u} \left(\pi(y) - q^{(m-1)}(v,y)\right) \left(\frac{q^{(1)}(y,u) - \pi(u)}{\pi(u)}\right) \leq \frac{1}{14}.
	\end{equation*}

	Plugging this in to \eqref{eqn: tv_to_uniform_distance_1} and taking maximum on both sides, we obtain
	\begin{equation}
	d_{\textrm{unif}}(q_v^{(m)}, \pi) \leq \frac{2}{7} + \frac{d_{\textrm{unif}}(q_v^{(m - 1)}, \pi)}{2} \label{eqn: tv_to_uniform_distance_2}
	\end{equation}
	The recurrence relation of \eqref{eqn: tv_to_uniform_distance_2} along with the initial condition $d_{\textrm{unif}}(q_v^{(t)}, \pi) \leq 1$ yields
	\begin{equation*}
	d_{\textrm{unif}}(q_v^{(t + 1)}, \pi) \leq \frac{8}{14} \Rightarrow d_{\textrm{unif}}(q_v^{(t + 2)}, \pi) \leq \frac{10}{28} \Rightarrow  d_{\textrm{unif}}(q_v^{(t + 3)}, \pi) \leq \frac{1}{4}
	\end{equation*}
	and the claim is shown.
\end{proof}

Together, Proposition \ref{prop: tv_mixing_time} and Lemma \ref{lem: tv_to_uniform_distance} imply the main result of this section.

\begin{proposition}
	\label{prop: pointwise_mixing_time}
	Assume $\min_{u \in v} \deg(u; G) \geq 10$. Then,
	\begin{equation*}
	\tau_{\infty}(G) \leq \frac{2}{\Phi^2(G)} \log \left(\frac{1440}{s(G)}\right)\log \left(\frac{14}{s(G)}\right)  + 3 \log \left(\frac{14}{s(G)}\right) + 3
	\end{equation*}
\end{proposition}
\begin{proof}
	Fix $a = \frac{1}{18}$, and let $\tau_0 = \frac{2}{\Phi^2(G)} \log \left(\frac{80}{a s(G)}\right)$. Then,
	\begin{equation*}
	\left(1 - \frac{\Phi^2(G)}{2}\right)^{\tau_0} \leq \exp(- \tau_0\Phi^2(G)/2) \leq \frac{as(G)}{80}
	\end{equation*}
	and so by Proposition \ref{prop: tv_mixing_time},
	\begin{equation*}
	\norm{q_v^{(\tau_0 + 3)} - \pi}_{TV} \leq \max \set{\frac{1}{20}, \frac{1}{8} + \frac{1}{20} + \frac{1}{20}} + \left(\frac{2}{as(G)}\right) \frac{as(G)}{80} \leq \frac{1}{4}.
	\end{equation*}

	It is a well known fact \cite{montenegro2002} that if $\norm{q_v^{(t)} - \pi}_{TV} \leq \frac{1}{4}$, then for any $) < \epsilon < 1$, $\norm{q_v^{(t\log_2(1/\epsilon))} - \pi}_{TV} \leq \epsilon$. Therefore, letting $\tau_1 = (\tau_0 + 3) \log(\frac{14}{s(G)})$, 
	\begin{equation*}
	\norm{q_v^{(\tau_1)} - \pi}_{TV} \leq \frac{1}{14 s(G)}
	\end{equation*}
	and so $d_{\textrm{unif}}(q_v^{(\tau_1 + 3)}, \pi) \leq \frac{1}{4}$. 
\end{proof}

\subsection{Proof of Proposition \ref{prop: tv_mixing_time}.}
\label{sec: proof_of_proposition_tv_mixing_time}

For arbitrary starting distribution $q$ (meaning $\textrm{supp}(q) \subseteq V$ and $\sum_{u \in V}q(u) = 1$), and for $t \geq 0$ an integer, let $q^{(t)}$ be the $t$-step probability distribution of the lazy random walk with starting distribution $q$. Consider the distance function $h_q^{(t)}, t \geq 0$, 
\begin{equation*}
h_q^{(t)}(x) = \max \left\{ \sum_{u \in V} \left(q^{(m)}(u) - \pi(u)\right)w(u)\right\}
\end{equation*}
where the maximum is over all $w: V \to [0,1]$ such that $0 \leq w(u) \leq 1$ for all $u$, and $\sum_{u \in V} w(u) \pi(u) = x$. Writing $h_v^{(t)}(x) := h_{e_v}^{(t)}(x)$ in a small abuse of notation, in Lemmas \ref{lem: lt_ub} and \ref{lem: mixing_time_small_sets}, we give an upper bound on $h_{e_v}^{(t)}(x)$ for all $0 \leq x \leq 1$.
